---
title: "Untitled"
output: html_document
---

Chapter 2

```{r}

pacman::p_load(tidyverse, fpp2, GGally)

```

2.1

```{r}

# time-series object
y <- ts(c(123, 39, 78, 52, 110),
        start = 2012)

```

Some visualization

```{r}

# load time-series data.
d <- melsyd

# autoplot - pretty funky. 
autoplot(d[, "Economy.Class"]) +
  labs(title = "Economy class Melbourne-Sydney",
       y = "Thousands",
       x = "Year")

# another plot
autoplot(a10) +
  ggtitle("Antidiabetic drug sales") +
  ylab("$ million") +
  xlab("Year")

# first good plot
ggseasonplot(a10, year.labels=TRUE, year.labels.left=TRUE) +
  ylab("$ million") +
  ggtitle("Seasonal plot: antidiabetic drug sales")

# don't really like these polar ones..
ggseasonplot(a10, polar=TRUE) +
  ylab("$ million") +
  ggtitle("Polar seasonal plot: antidiabetic drug sales") 

# subseries plot
ggsubseriesplot(a10) +
  ylab("$ million") +
  ggtitle("Seasonal subseries plot: antidiabetic drug sales")

# line-plot
autoplot(elecdemand[,c("Demand", "Temperature")], 
         facets=TRUE) +
  xlab("Year: 2014") + ylab("") +
  ggtitle("Half-hourly electricity demand: Victoria, Australia")

# fucking qplot
qplot(Temperature, Demand, data=as.data.frame(elecdemand)) +
  ylab("Demand (GW)") + xlab("Temperature (Celsius)")

# facet
autoplot(visnights[,1:5], facets=TRUE) +
  ylab("Number of visitor nights each quarter (millions)")

# pairs plot
GGally::ggpairs(as.data.frame(visnights[,1:5]))

# lag plots
# clearly lag 8 is pretty strong, seasonal?? 
beer2 <- window(ausbeer, start=1992)
gglagplot(beer2)

# autocorrelation
ggAcf(beer2)

# trend & seasonality
aelec <- window(elec, start=1980)
autoplot(aelec) + xlab("Year") + ylab("GWh")
ggAcf(aelec, lag=48)

# white noise
set.seed(30)
y <- ts(rnorm(50))
autoplot(y) + ggtitle("White noise")
ggAcf(y)

```

# Exercises

## E1
Use the help function to explore what the series gold, woolyrnq and gas represent.

Use autoplot() to plot each of these in separate plots.
What is the frequency of each series? Hint: apply the frequency() function.

Use which.max() to spot the outlier in the gold series. Which observation was it?

```{r}

# daily morning gold priors in US dollars (1985-1989)
?gold

# quarterly prod. of woollwn yarin in AUS (1965-1994)
?woolyrnq

# AUS monthly gas production (1956-1995)
?gas

# plot and frequency
plotfreq <- function(d){
  
  plt <- d %>%
    autoplot()
  
  print(frequency(d))
  
  return(plt)
  
}

gold_plot <- plotfreq(gold)
wool_plot <- plotfreq(woolyrnq)
gas_plot <- plotfreq(gas)

cowplot::plot_grid(gold_plot, wool_plot, gas_plot)

# outlier in gold: 
which.max(gold)

```

## E2
need files for the rest of the exercises..

